### 2024-03-13 outputs/paper.qmd
initial version, not finished yet

---
title: "Evaluating the Effectiveness of Statistical Models in Predicting Mortality RatesAn Analysis of Negative Binomial Distribution Model in Canada"
subtitle: "An Analysis of Negative Binomial Distribution Model"
author: Xiyou Wang, Yetao Guo
thanks: "Code and data are available at: https://github.com/wxywxy666/Mortality-in-Canada."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "NOT FINISHED YET!."
output:
  bookdown::pdf_document2:
    toc-title: "Contents"
toc: true
toc-title: "Contents"
number-sections: true
format: pdf
bibliography: references.bib
---
```{r setup, include=FALSE}
# Load libraries
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
```

\newpage
# Introduction
Understanding the field of public health mortality rate is important for policymakers, healthcare professionals, and society in general. The barometers of the mortality rate serve the efficiency of the well-being of society and the health system. This provides more knowledge on the challenges of general health, thus guiding resource allocation. A key case study emerges for Canada with the reflection of the morality rate landscape through the more overall societal trends and the dynamics of healthcare. In this research study, the morality complicated rate is solved within the population of Canada through the age-specific dissection of mortality rate and the lead on the causes of death illustration in the various cohorts demographically; thus, our main aim is to give more insights on the factors underlying shaping Canada's mortality trends.

Due to COVID-19, The number of casualties in this research is the key critical revolution. At the same time, it acts as the subsequent analysis and the strategies for intervention. While the research findings are highlighted in the abstract, the introduction explores the comprehensive understanding of the research, providing an overview that is extensive while on an exhaustive component, ensuring there is refrainment. This research study sets a foundation for strategic address and decision-making based on evidence. Additionally, we seek a better understanding of the efforts guiding an intervention that is subsequent through the research study. To ensure the research findings efforts for the future have a structured analysis are presented in the section following an analysis that is more structured exploring the importance of Canada's results on the employed methodologies. Through the research structure description, the main aim is to offer the readers an informed road map for the multifaceted landscape navigation concerning the Canadian mortality rate.

\newpage
# Measurement/Methodology
The dataset comes from Statistics Canada and focuses on mortality trends in Canada from 2000 to 2022, categorized by age, sex, and cause of death. This information is crucial for understanding public health trends and guiding policy decisions. Therefore, this dataset was selected due to its comprehensive coverage, high data quality and reliability.
Data is updated annually, and the specific data used in this article is the latest available as of January 27, 2023. Raw data set presents data on the total number of deaths in Canada and ranks the leading causes of death, such as salmonella infections, shigellosis and amoebiasis, and tuberculosis, etc.

Also included are maps of the age at time of death, the distribution of both genders, and partial places of residence. It is worth noting that The category "Age at time of death,  all ages" includes the number of deaths for children aged under one year old,The deaths for which age is not stated are  included in the "Age at the time of death, all ages" category but not distributed among age groups.
All of the data is possessed and cleaned through R studio,a programming language for statistical computing and graphics.

\newpage
# Data：
This section seeks to provide a comprehensive understanding of the dataset used in our analysis. The dataset captures the mortality rates across various demographics in Canada from 2000 to 2022. The data offer a wider perspective allowing trends to be analyzed over time, including periods of COVID-19 outbreaks.

## raw data
First, the original dataset file of "Leading causes of death", named "1310039401-eng", was downloaded from Statistics Canada(https://www.statcan.gc.ca). Five key variables are included. “Leading causes of death” tabulated is the underlying cause of death. This is defined as the disease or injury which initiated the train of events leading directly to death, or the circumstances of the accident or violence which produced the fatal injury. is selected from the conditions listed on the medical certificate of cause of death. "Characteristics" contains the "Rank of leading causes of death" which is based on the "Number of deaths". "Age at time of death" is attained at the last birthday preceding death. "Reference period" contains specific numbers, from 2000 to 2022. This article mainly studies the statistical model of death toll, so only the data of all ages and both sexes has been downloaded.

## data cleaning
```{r echo = FALSE}
# Load cleaned data
cleaned_data <- read.csv(here::here("inputs/cleaned_data"))
cleaned_data <- cleaned_data[, -1]

# Create the filtered and mutated table
cleaned_data |>
  filter(
    year == 2022,
    ranking <= 10
  ) |>
  mutate(total_deaths = format(total_deaths, big.mark = ",")) |>
  kable(
    col.names = c("Year", "Cause", "Deaths", "Ranking", "Years"),
    align = c("l", "r", "r", "r", "r"),
    digits = 0, 
    booktabs = TRUE, 
    linesep = ""
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
In order to understand the data more intuitively, data cleansing is necessary. A detailed description of each variable explains how these variables are critical for understanding public health trends and policy implications.

table1 showcases the first ten rows of Leading Causes of Death in Canada for the Year 2022.To further understand the table,"cause" refers to the event that led to the deaths recorded in the dataset. Each cause is identified by its medical name."deaths" represents the total number of people who passed away due to each listed cause. Years represent..........It should be noted that since the COVID-19 outbreak began in 2020, the data for 'years' is limited to only three years

\newpage
# Result
```{r echo = FALSE}
# Select top five causes
top_five <-
  cleaned_data |>
  filter(
    year == 2022,
    n == 23
  )
ca_cod_top_five <- top_five$cause[1:5]

cleaned_data <-
  cleaned_data |>
  filter(cause %in% ca_cod_top_five)

# Using ggplot to make figure
cleaned_data |>
  ggplot(aes(x = year, y = total_deaths, color = cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Annual number of deaths in Canada") +
  facet_wrap(vars(cause), scales = "free_y", dir = "v", ncol = 1) +
  theme(legend.position = "none")
```
\newpage
# Discussion：
Plots of time-series changes in the number of deaths in Canada due to different factors are provided in the visualisation results of this study. The excellence of two forecasting models for mortality is analysed in comparison.

## Finding
The main focus of this document revolves around a comparison of the performance of two statistical models - the negative binomial distribution model and the Poisson distribution model - in predicting mortality in Canada. Through detailed data analysis, we find that the negative binomial distribution model outperforms the Poisson distribution model in terms of both the expected log-point-by-point predicted density difference (elpd_diff) and the k-fold cross-validated ELPD value (elpd_kfold). In addition, the negative binomial distribution model has a smaller standard deviation, suggesting that its predictions are more stable. This finding provides a new perspective that negative binomial distribution models may be more appropriate for describing and predicting mortality in Canada.

## Economic Impact Insigths
From the perspective of economic impact, accurate mortality prediction is an important reference value for a number of industries, including insurance, pension, and healthcare. The superiority of the negative binomial distribution model means that we can predict future mortality rates more accurately, thus helping these industries to make more rational economic decisions, such as setting insurance premium rates and planning medical resources. In addition, for governments, accurate mortality prediction can also help to formulate more effective social security policies and reduce financial pressure. In terms of insight, the reason why the negative binomial distribution model can achieve better prediction results may be related to its ability to better handle the discrete and over-discrete nature of the data. This suggests that we need to pay more attention to the characteristics of the data and choose more appropriate statistical models when dealing with similar problems.

## Societal and Technological Influences
In terms of social impact, accurate mortality forecasts help the public to better understand the trends in demographic and health conditions, and thus make more informed life decisions. For example, the public can adjust their health management and retirement planning based on the prediction results. In terms of technical implications, the findings in this paper provide new ideas for the application of statistical models in the field of mortality prediction. In the future, we can further explore and optimise the negative binomial distribution model, or find other more suitable models to improve the prediction accuracy and stability.

## Weakness and Future Research Directions
Despite the superiority of the negative binomial distribution model in predicting mortality in Alberta, there are still some shortcomings. For example, only two models were compared in this paper and other possible models were not considered; furthermore, the parameters of the models were not analysed in detail to identify the key factors affecting the prediction results. Future research directions can be developed in the following aspects: firstly, other possible statistical models can be further explored to find a more suitable model for predicting mortality in the Canada; secondly, an in-depth analysis of the model's parameters can be carried out to reveal the specific factors affecting the prediction effect; and lastly, attempts can be made to incorporate more factors into the model in order to improve the prediction accuracy and the scope of application. In summary, the findings of this paper are of great guiding significance for the modelling of mortality in the Canada, and also provide us with an opportunity to explore in depth the application of statistical models in the field of mortality prediction.

\newpage
# Conclusion

### 2024-03-13 scripts/02-data_cleaning.R

@@ -1,0 +1,55 @@
+# set up
+library(dplyr)
+
+# Read the CSV file as lines
+lines <- readLines("inputs/1310039401-eng.csv")
+
+# Skip specified lines
+lines_to_keep <- lines[-c(1:11, 13:15, 118:152)]
+
+data <- read.csv(text = lines_to_keep, check.names = FALSE)
+
+# roughly clean data
+for (i in 2:nrow(data)) {
+  if (i %% 2 == 0) {
+    data[i, 1] <- data[i - 1, 1]
+  }
+}
+
+data <- data[seq(0,nrow(data),2),]
+data <- data[, -2]
+names(data)[1] <- "cause"
+
+# totally change the form into a new data frame
+new <- data.frame(year = numeric(), cause = character(), total_deaths = numeric())
+
+for (i in 1:nrow(data)) {
+  for (n in 2:ncol(data)) {
+    if (data[i, n] == "..")
+      data[i, n] <- 0
+    if (data[i, n] != 0) {
+      new <- rbind(new, data.frame(year = as.numeric(names(data)[n]),
+                                   cause = data$cause[i],
+                                   total_deaths = data[i, n]))
+    }
+  }
+}
+
+# clean cause names
+for (i in 1:length(new$cause)) {
+  new$cause[i] <- gsub("\\s*\\[.*?\\]", "", new$cause[i])
+}
+
+# sort and add ranking
+new <- new %>%
+  arrange(year, desc(as.numeric(gsub(",", "", total_deaths))))%>%
+  group_by(year) %>%
+  mutate(ranking = row_number())
+
+# count times of appear
+new <- new %>%
+  mutate(total_deaths = as.numeric(gsub(",", "", total_deaths))) %>%
+  group_by(cause) %>%
+  mutate(n = sum(total_deaths != 0))
+
+write.csv(new, "inputs/cleaned_data")

### 2024-03-14 outputs/paper.qmd
context upgrade, mainly on data and result

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "NOT FINISHED YET!."
+abstract: "This study explores the complex relationship between Canadian social trends, healthcare dynamics, and mortality rates.
+We compared the goodness of fit between the negative binomial distribution model and the Poisson distribution model using a comprehensive dataset, and conducted in-depth research on the complexity of mortality rates.
+Our main findings reveal important associations between various social and healthcare factors and mortality rates, indicating that the negative binomial distribution model is more suitable for the data.
+Our research contributes to a deeper understanding of mortality rates and their influencing factors, laying a valuable foundation for future research and strategic initiatives aimed at improving public health outcomes."
 output:
@@
 # Introduction
-Understanding the field of public health mortality rate is important for policymakers, healthcare professionals, and society in general.
-The barometers of the mortality rate serve the efficiency of the well-being of society and the health system.
-This provides more knowledge on the challenges of general health, thus guiding resource allocation.
-A key case study emerges for Canada with the reflection of the morality rate landscape through the more overall societal trends and the dynamics of healthcare.
-In this research study, the morality complicated rate is solved within the population of Canada through the age-specific dissection of mortality rate and the lead on the causes of death illustration in the various cohorts demographically; thus, our main aim is to give more insights on the factors underlying shaping Canada's mortality trends.
+The field of public health mortality is of great significance to policy makers, healthcare professionals, and society as a whole.
+Mortality rate is an important indicator for measuring social well-being and health system efficiency.
+They provide valuable insights into the challenges faced by understanding the overall health situation, thereby providing guidance for resource allocation and policy decision-making.
 
-Due to COVID-19, The number of casualties in this research is the key critical revolution.
-At the same time, it acts as the subsequent analysis and the strategies for intervention.
-While the research findings are highlighted in the abstract, the introduction explores the comprehensive understanding of the research, providing an overview that is extensive while on an exhaustive component, ensuring there is refrainment.
-This research study sets a foundation for strategic address and decision-making based on evidence.
-Additionally, we seek a better understanding of the efforts guiding an intervention that is subsequent through the research study.
-To ensure the research findings efforts for the future have a structured analysis are presented in the section following an analysis that is more structured exploring the importance of Canada's results on the employed methodologies.
-Through the research structure description, the main aim is to offer the readers an informed road map for the multifaceted landscape navigation concerning the Canadian mortality rate.
+In this study, we delved into the complexity of mortality rates, particularly in the context of Canada.
+Through a comprehensive analysis of the mortality rate situation, we aim to reflect on Canada's overall social trends and healthcare dynamics.
+Our study provides a subtle understanding of mortality, highlighting its complexity and influencing factors.
 
-\newpage
-# Measurement/Methodology
-The dataset comes from Statistics Canada and focuses on mortality trends in Canada from 2000 to 2022, categorized by age, sex, and cause of death.
-This information is crucial for understanding public health trends and guiding policy decisions.
-Therefore, this dataset was selected due to its comprehensive coverage, high data quality and reliability.
-Data is updated annually, and the specific data used in this article is the latest available as of January 27, 2023.
-Raw data set presents data on the total number of deaths in Canada and ranks the leading causes of death, such as salmonella infections, shigellosis and amoebiasis, and tuberculosis, etc.
+In this study, our primary estimand is the mortality  in Canada from 2000 to 2022, by top cause each year.
+We perform in-depth analysis by fitting mathematical models, taking into account the impact of various social and health care factors on mortality.
 
-Also included are maps of the age at time of death, the distribution of both genders, and partial places of residence.
-It is worth noting that The category "Age at time of death,  all ages" includes the number of deaths for children aged under one year old,The deaths for which age is not stated are  included in the "Age at the time of death, all ages" category but not distributed among age groups.
-All of the data is possessed and cleaned through R studio,a programming language for statistical computing and graphics.
+One important result of our analysis is that there is a complex relationship between mortality rate and various social and healthcare factors.
+This discovery is the focus of our report, providing valuable insights into the challenges and opportunities facing the Canadian healthcare system.
+
+We have provided more details in this introduction than in the abstract, but we will not disclose the full content of our research findings.
+On the contrary, we have provided a high-level overview of our research findings, laying the foundation for subsequent chapters of this article.
+
+Looking ahead, we discussed the next steps of our research and outlined the direction for future exploration.
+Finally, we conclude the introduction with a brief paragraph, emphasizing the structure of the paper and guiding readers to understand the following chapters and their respective contributions to the overall narrative.
+
+By exploring the complexity of Canada's public health mortality rate, we hope to deepen our understanding of this key indicator and its impact on policy-making, healthcare services, and social well-being.
 
@@
 
-## raw data
+## Raw data
 First, the original dataset file of "Leading causes of death", named "1310039401-eng", was downloaded from Statistics Canada(https://www.statcan.gc.ca).
@@
 
-## data cleaning
+## Data cleaning
+In order to understand the data more intuitively, data cleansing is necessary.
+A detailed description of each variable explains how these variables are critical for understanding public health trends and policy implications.
+
+table1 showcases the first ten rows of Leading Causes of Death in Canada for the Year 2022.To further understand the table,"cause" refers to the event that led to the deaths recorded in the dataset.
+Each cause is identified by its medical name."deaths" represents the total number of people who passed away due to each listed cause.
+Years represent..........It should be noted that since the COVID-19 outbreak began in 2020, the data for 'years' is limited to only three years
+
+### Preview
 ```{r echo = FALSE}
@@
 ```
-In order to understand the data more intuitively, data cleansing is necessary.
-A detailed description of each variable explains how these variables are critical for understanding public health trends and policy implications.
+table1 showcases the first ten rows of Leading Causes of Death in Canada for the Year 2022.
+To further understand the table, "Cause" refers to the event that led to the deaths recorded in the dataset.
+Each cause is identified by its medical name."Deaths" represents the total number of people who passed away due to each listed cause.
+“Years” represent..........It should be noted that since the COVID-19 outbreak began in 2020, the data for 'years' is limited to only three years
 
-table1 showcases the first ten rows of Leading Causes of Death in Canada for the Year 2022.To further understand the table,"cause" refers to the event that led to the deaths recorded in the dataset.
-Each cause is identified by its medical name."deaths" represents the total number of people who passed away due to each listed cause.
-Years represent..........It should be noted that since the COVID-19 outbreak began in 2020, the data for 'years' is limited to only three years
-
-\newpage
-# Result
+### Visualization
+Since the entire data set is so large, a simple approach would be to focus on the top five causes of death.
+Of course, these five causes of death must occur every year from 2000 to 2022, which means that its "Years" value is 23.
 ```{r echo = FALSE}
@@
 \newpage
+# Result
+This table tells us the summary statistics of the number of yearly deaths by cause in Canada.
+The number of observations in the data is 115.
+The minimum value in the data set is 8521 and the maximum value is 82822.SD refers to Stands for Standard Deviation, a measure of how spread out numbers are in the set is 25849.99.
+Var refers to variance, indicating how much the numbers vary from the average, and it's 668221779.
+
+## Model fitting
+The Poisson distribution describes the probability distribution of the number of events that occur within a fixed unit of time or space. Its probability mass function is:
+$$P(X = k) = \frac{e^{-\lambda}\lambda^k}{k!}$$
+The Poisson distribution is usually more concise and convenient, can fit a wider range of models and is easier to interpret.
+Nevertheless, Poisson distribution assumes that the mean($\lambda$) and variance($\lambda$) are equal.
+
+Compared with the Poisson distribution, the calculation and derivation of the negative binomial distribution are more complicated, and the parameter estimation of the model may be more difficult, especially when the amount of data is small or the model does not fit well.
+Its probability mass function is:
+$$P(X = k) = \binom{k+r-1}{k}p^r(1-p)^{k-r}$$
+However, the negative binomial distribution is a generalization of the Poisson distribution that allows for differences between mean($\frac{r(1-p)}{p}$) and variance($\frac{r(1-p)}{p^2}$).
+It does this by processing an extra parameter: the dispersion parameter($r$), which is the number of successes that need to be observed before a number($k$) of observed failures.
+
+When means and variances are not equal, especially when overdispersion occurs, using the negative binomial distribution instead of the Poisson distribution can provide a more accurate statistical model that better fits the data.
+For the sake of rigor, this study still completely compares the two.
+
+## Estimate
+Implementing poisson and negative binomial regression, simultaneously.
+Model the most prevalent cause of deaths.
+According to the table, the estimates are similar, so it is necessary to use a posterior predictive check in addition.
+
+## Posterior predictive check
+There are many observations with a pareto_k > 0.7.
+Large Pareto's K means the model posterior would be too different if one data point is being removed.
+This suggests that the model is not capturing the data well (i.e.
+some data points with high K are highly influential and not being considered by the model).
+
+## Resampling
+Following this logic, LOOCV(Leave-one-out cross-validation) will not be trustable anymore if most of the Pareto's K is higher than 0.7, although it can usually get a more accurate prediction.
+With this many problematic observations, K-fold cross-validation with argument 'K=10' to perform 10-fold cross-validation, that the data used for each training becomes closer to the entire data set, and the deviation decreases.
+
+It is less computationally intensive, and more efficient when processing large data sets.
+So K-fold will be a more suitable resampling method in this study.
+
+The information provided in the table allows us to compare the relative performance of the negative binomial distribution model and the Poisson distribution model.
+The result mainly relies on the “elpd_diff” and “elpd_kfold” values of the two models, namely "expected log pointwise predictive density difference" and "ELPD for K-fold cross-validation".
+The higher these two values are, the better the model's predictive performance is.
+
+From table, we can see that the negative binomial distribution model has higher “elpd_diff” and “elpd_kfold” values and also has a smaller standard deviation.
+This shows that the negative binomial distribution model has better predictive performance than the Poisson distribution model in cross-validation.
+Therefore, we can infer that the negative binomial distribution model is a better fit for the overall number of deaths by cause, in Canada, from 2000 to 2022.
+
+\newpage
 # Discussion：
@@
 The excellence of two forecasting models for mortality is analysed in comparison.
-
 ## Finding
@@
 
-## Economic Impact Insigths
+## Economic Impact Insights
 From the perspective of economic impact, accurate mortality prediction is an important reference value for a number of industries, including insurance, pension, and healthcare.

### 2024-03-17 scripts/00-data_simulating.R
finish readme

@@ -1,0 +1,22 @@
+# basic set
+set.seed(888)
+num_entries <- 32
+
+# create year, cause, total_death and years
+years <- sample(2000:2022, num_entries, replace = TRUE)
+causes <- sample(c("Malignant neoplasms", "Diseases of heart", 
+                   "Cerebrovascular diseases", "Chronic lower respiratory diseases",
+                   "Accidents (unintentional injuries)", "Diabetes mellitus", 
+                   "Influenza and pneumonia", "Alzheimer’s disease",
+                   "Nephritis, nephrotic syndrome, and nephrosis", "Septicemia"),
+                 num_entries, replace = TRUE)
+total_deaths <- sample(5000:70000, num_entries, replace = TRUE)
+years_column <- ifelse(runif(num_entries) < 0.9, 23, sample(1:22, num_entries, replace = TRUE))
+
+# make it a frame
+data <- data.frame(year = years, cause = causes, total_deaths = total_deaths, years = years_column)
+
+print(data)
+
+# save simulated data
+write.csv(data, "inputs/simulated_data")

### 2024-03-17 scripts/01-data_test.R

@@ -1,0 +1,20 @@
+# read simualted data
+data <- read.csv("inputs/simulated_data")
+
+# create function to test
+test_data <- function(data) {
+  year_test <- all(sapply(data$year, is.numeric) & data$year >= 2000 & data$year <= 2022)
+  years_test <- all(sapply(data$years, is.numeric) & data$years >= 0 & data$years <= 23)
+  cause_test <- all(sapply(data$cause, is.character))
+  total_deaths_test <- all(sapply(data$total_deaths, is.numeric) & data$total_deaths > 0)
+  
+  return(list(
+    year_test = year_test,
+    years_test = years_test,
+    cause_test = cause_test,
+    total_deaths_test = total_deaths_test
+  ))
+}
+
+# run test function
+test_data(data)

### 2024-03-17 scripts/02-data_cleaning.R

@@ -50,6 +50,6 @@ new <- new %>%
 new <- new %>%
   mutate(total_deaths = as.numeric(gsub(",", "", total_deaths))) %>%
   group_by(cause) %>%
-  mutate(n = sum(total_deaths != 0))
+  mutate(years = sum(total_deaths != 0))
 
 write.csv(new, "inputs/cleaned_data")

### 2024-03-17 outputs/paper.qmd
1

@@
 ---
-title: "Evaluating the Effectiveness of Statistical Models in Predicting Mortality RatesAn Analysis of Negative Binomial Distribution Model in Canada"
-subtitle: "An Analysis of Negative Binomial Distribution Model"
+title: "Evaluating the Effectiveness of Statistical Models in Analyzing Mortality in Canada"
+subtitle: "A Practical Application of Negative Binomial Regression"
 author: Xiyou Wang, Yetao Guo
@@
 abstract: "This study explores the complex relationship between Canadian social trends, healthcare dynamics, and mortality rates.
-We compared the goodness of fit between the negative binomial distribution model and the Poisson distribution model using a comprehensive dataset, and conducted in-depth research on the complexity of mortality rates.
-Our main findings reveal important associations between various social and healthcare factors and mortality rates, indicating that the negative binomial distribution model is more suitable for the data.
+We compare the differences between Poisson and negative binomial distribution models using a comprehensive dataset and provide insights into the complexity of mortality.
+Our main findings reveal important associations between various social and healthcare factors and mortality rates, indicating that the negative binomial regression is more suitable for the data.
 Our research contributes to a deeper understanding of mortality rates and their influencing factors, laying a valuable foundation for future research and strategic initiatives aimed at improving public health outcomes."
@@
 library(kableExtra)
+library(modelsummary)
+library(rstanarm)
+library(broom.mixed)
+library(loo)
+
+options(mc.cores = parallel::detectCores())
+rstan::rstan_options(auto_write = TRUE)
+options(rstan.verbose = FALSE)
 ```
-
 \newpage
 # Introduction
-The field of public health mortality is of great significance to policy makers, healthcare professionals, and society as a whole.
-Mortality rate is an important indicator for measuring social well-being and health system efficiency.
-They provide valuable insights into the challenges faced by understanding the overall health situation, thereby providing guidance for resource allocation and policy decision-making.
+Studying public health mortality rates is of great significance to policy makers, healthcare professionals, and society as a whole.
+Mortality is an important indicator of social well-being and health system efficiency.
+It provides valuable insights for addressing the challenges faced by understanding the overall health situation, which in turn guides resource allocation and policy decisions.
 
-In this study, we delved into the complexity of mortality rates, particularly in the context of Canada.
-Through a comprehensive analysis of the mortality rate situation, we aim to reflect on Canada's overall social trends and healthcare dynamics.
-Our study provides a subtle understanding of mortality, highlighting its complexity and influencing factors.
-
-In this study, our primary estimand is the mortality  in Canada from 2000 to 2022, by top cause each year.
-We perform in-depth analysis by fitting mathematical models, taking into account the impact of various social and health care factors on mortality.
+For the purposes of this study, our primary estimate is mortality in Canada from 2000 to 2022, broken down by major cause for each year.
+In-depth analyses were conducted by fitting mathematical models, taking into account the impact of various social and health care factors on mortality.
+Through a comprehensive analysis of the mortality situation, we aim to reflect on overall social trends and health care dynamics in Canada.
+Our study provides a nuanced understanding of mortality, highlighting its complexity and the factors that influence it.
 
@@
 This discovery is the focus of our report, providing valuable insights into the challenges and opportunities facing the Canadian healthcare system.
+By exploring the complexity of Canada's public health mortality rate, we hope to deepen our understanding of this key indicator and its impact on policy-making, healthcare services, and social well-being.
 
-We have provided more details in this introduction than in the abstract, but we will not disclose the full content of our research findings.
-On the contrary, we have provided a high-level overview of our research findings, laying the foundation for subsequent chapters of this article.
-
-Looking ahead, we discussed the next steps of our research and outlined the direction for future exploration.
-Finally, we conclude the introduction with a brief paragraph, emphasizing the structure of the paper and guiding readers to understand the following chapters and their respective contributions to the overall narrative.
-
-By exploring the complexity of Canada's public health mortality rate, we hope to deepen our understanding of this key indicator and its impact on policy-making, healthcare services, and social well-being.
+We use R[@R] for all data wrangling and analysis and R packages dplyr[@dplyr] to clean the original data, ggplot2[@ggplot2], knitr[@knitr], kableExtra[@kableExtra] to produce the charts and modelsummary[@modelsummary], rstanarm[@rstanarm], broom.mixed[@broom], loo[@loo] to fit the data into the model and generate relevant analysis charts.
 
@@
 This section seeks to provide a comprehensive understanding of the dataset used in our analysis.
-The dataset captures the mortality rates across various demographics in Canada from 2000 to 2022.
 The data offer a wider perspective allowing trends to be analyzed over time, including periods of COVID-19 outbreaks.
 
+## Source
+Our study uses a dataset from Statistics Canada[@statcan] and focuses on mortality trends in Canada from 2000 to 2022, categorized by age, sex, and cause of death.
+This information is crucial for understanding public health trends and guiding policy decisions.
+Therefore, this dataset was selected due to its comprehensive coverage, high data quality and reliability.
+
+Data from Statistics Canada is updated annually, and the specific data used in this article is the latest available as of 2022 .
+Raw data set presents data on the total number of deaths in Canada and ranks the leading causes of death, such as salmonella infections, shigellosis and amoebiasis, and tuberculosis, etc.
+Also included are maps of the age at time of death, the distribution of both genders, and partial places of residence.
+It is worth noting that The category "Age at time of death, all ages" includes the number of deaths for children aged under one year old, The deaths for which age is not stated are included in the "Age at the time of death, all ages" category but not distributed among age groups.
+All of the data is possessed and cleaned through R studio, a programming language for statistical computing and graphics.
+
+### Measurement
+Using Poisson and negative binomial distributions as link functions, the data were fitted and the results of the two distributions were compared to select the best link function.
+The Poisson distribution is a simple way to predict events, but it expects the average number of events to be the same as the amount they vary, which isn't always true for actual data.The negative binomial distribution is a more flexible version that can deal with differences between the average number of events and how much they vary by including an extra detail, the dispersion parameter, that helps when the data is more spread out than the average.
+
 ## Raw data
-First, the original dataset file of "Leading causes of death", named "1310039401-eng", was downloaded from Statistics Canada(https://www.statcan.gc.ca).
+First, the original dataset file of "Leading causes of death", named "1310039401-eng", was downloaded from Statistics Canada.
 Five key variables are included.
-“Leading causes of death” tabulated is the underlying cause of death.
-This is defined as the disease or injury which initiated the train of events leading directly to death, or the circumstances of the accident or violence which produced the fatal injury.
-is selected from the conditions listed on the medical certificate of cause of death.
+"Leading cause of death" in the table is defined as an illness or injury that triggers a sequence of events leading directly to death, or an accident or violent situation that results in a fatal injury.
+The underlying cause is selected from the conditions listed on the medical certificate of cause of death.
 "Characteristics" contains the "Rank of leading causes of death" which is based on the "Number of deaths".
@@
 "Reference period" contains specific numbers, from 2000 to 2022.
-This article mainly studies the statistical model of death toll, so only the data of all ages and both sexes has been downloaded.
 
@@
 In order to understand the data more intuitively, data cleansing is necessary.
-A detailed description of each variable explains how these variables are critical for understanding public health trends and policy implications.
-
-table1 showcases the first ten rows of Leading Causes of Death in Canada for the Year 2022.To further understand the table,"cause" refers to the event that led to the deaths recorded in the dataset.
-Each cause is identified by its medical name."deaths" represents the total number of people who passed away due to each listed cause.
-Years represent..........It should be noted that since the COVID-19 outbreak began in 2020, the data for 'years' is limited to only three years
+The other reason is that the original data file contains many irrelevant instructions and variables.
+The first step is to delete them directly and to rename “Leading causes of death” as “cause”.
+And then, replace the original classification method based on the cause of death with year, named “year”.
+Next, add two variables, "ranking" represents the ranking of the number of people for this cause in the same year, and "years" represents the number of times this cause appears from 2000 to 2022.
+Finally, sort them with "year" as the main one, and then sort them in the reverse order of "total_death" according to this cause in this year, that is, in the forward order of "ranking".
 
@@
 ```{r echo = FALSE}
-# Load cleaned data
-cleaned_data <- read.csv(here::here("inputs/cleaned_data"))
+#| label: tbl-1
+#| tbl-cap: "Modeling the most prevalent cause of deaths in Canada, 2000-2022"
+cleaned_data <- read.csv("cleaned_data")
 cleaned_data <- cleaned_data[, -1]
@@
 ```
-table1 showcases the first ten rows of Leading Causes of Death in Canada for the Year 2022.
-To further understand the table, "Cause" refers to the event that led to the deaths recorded in the dataset.
-Each cause is identified by its medical name."Deaths" represents the total number of people who passed away due to each listed cause.
-“Years” represent..........It should be noted that since the COVID-19 outbreak began in 2020, the data for 'years' is limited to only three years
 
-### Visualization
-Since the entire data set is so large, a simple approach would be to focus on the top five causes of death.
-Of course, these five causes of death must occur every year from 2000 to 2022, which means that its "Years" value is 23.
+Preview the processed clean data first, which is very helpful for understanding the overall data.
+@tbl-1 shows the top-ten causes in 2022, from which some rough findings can be obtained.
+We previously predicted that the number of occurrences of most causes of death would be 23, which means they would occur every year, but the actual situation is that except for COVID-19, the rest are 23, and COVID-19 is 3, because COVID-19 suddenly appeared in 2020.
+Does this mean that Canada’s medical and health system has not undergone significant changes?
+The same diseases are still plaguing the people in 23 years.
+
+### Trend
 ```{r echo = FALSE}
+#| label: fig-1
+#| fig-cap: "Annual number of deaths for the top-five causes in 2022, since 2001, for Canada"
 # Select top five causes
@@
     year == 2022,
-    n == 23
+    years == 23
   )
@@
 ```
+
+@fig-1 shows how many people died from different causes each year in Canada from 2000 to around 2022.
+The vertical axis represents the annual number of deaths.
+The horizontal axis represents time, from the year 2000 to about 2024.
+Different colored lines for each cause show whether the number of deaths is going up, going down, or staying about the same over time.
+The red line has some fluctuations, but the overall trend is upward, indicating that deaths from accidents have been rising.
+The blue line is mostly flat with a bit of a drop, which could mean fewer people are dying from strokes and related diseases.
+This suggests improvements in healthcare or prevention measures for these conditions.
+The green line climbs up slowly, indicating more deaths from diseases like emphysema.
+The purple line indicates The number of deaths from heart disease is the highest among the listed causes and shows a slight upward trend over the two decades, with some fluctuations but no significant increase or decrease.
+Orange Line: The trend for cancer-related deaths is gradually increasing, indicating a growing number of deaths due to malignant neoplasms over the years.
+
 \newpage
 # Result
-This table tells us the summary statistics of the number of yearly deaths by cause in Canada.
+```{r echo = FALSE}
+#| label: tbl-2
+#| tbl-cap: "Summary statistics of the number of yearly deaths, by cause, in Canada"
+death <- data.frame(total <- as.numeric(cleaned_data$total_deaths))
+
+summary_table <- death %>%
+  summarise(
+    Min = min(total),
+    Mean = mean(total),
+    Max = max(total),
+    SD = sd(total),
+    Var = var(total),
+    N = n()
+  )
+
+# Create a styled table using kable and kableExtra
+kable(summary_table, table.attr = "style='width:100%;'") %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) %>%
+  row_spec(0, bold = T)
+```
+
+@tbl-2 tells us the summary statistics of the number of yearly deaths by cause in Canada.
 The number of observations in the data is 115.
@@
 ## Model fitting
-The Poisson distribution describes the probability distribution of the number of events that occur within a fixed unit of time or space. Its probability mass function is:
-$$P(X = k) = \frac{e^{-\lambda}\lambda^k}{k!}$$
-The Poisson distribution is usually more concise and convenient, can fit a wider range of models and is easier to interpret.
+The Poisson distribution describes the probability distribution of the number of events that occur within a fixed unit of time or space.
+Its probability mass function is: $$P(X = k) = \frac{e^{-\lambda}\lambda^k}{k!}$$ The Poisson distribution is usually more concise and convenient, can fit a wider range of models and is easier to interpret.
 Nevertheless, Poisson distribution assumes that the mean($\lambda$) and variance($\lambda$) are equal.
@@
 Compared with the Poisson distribution, the calculation and derivation of the negative binomial distribution are more complicated, and the parameter estimation of the model may be more difficult, especially when the amount of data is small or the model does not fit well.
-Its probability mass function is:
-$$P(X = k) = \binom{k+r-1}{k}p^r(1-p)^{k-r}$$
-However, the negative binomial distribution is a generalization of the Poisson distribution that allows for differences between mean($\frac{r(1-p)}{p}$) and variance($\frac{r(1-p)}{p^2}$).
+Its probability mass function is: $$P(X = k) = \binom{k+r-1}{k}(1-p)^{k}p^r$$ However, the negative binomial distribution is a generalization of the Poisson distribution that allows for differences between mean($\frac{r(1-p)}{p}$) and variance($\frac{r(1-p)}{p^2}$).
 It does this by processing an extra parameter: the dispersion parameter($r$), which is the number of successes that need to be observed before a number($k$) of observed failures.
@@
 ## Estimate
-Implementing poisson and negative binomial regression, simultaneously.
+Implementing Poisson and negative binomial regression, simultaneously.
 Model the most prevalent cause of deaths.
-According to the table, the estimates are similar, so it is necessary to use a posterior predictive check in addition.
+According to the @tbl-3, the estimates are similar, so it is necessary to use a posterior predictive check in addition.
+```{r  echo=FALSE, warning = FALSE}
+#| label: tbl-3
+#| tbl-cap: "Modeling the most prevalent cause of deaths in Canada, 2001-2020"
+cause_of_death_canada_poisson <-
+  stan_glm(
+    total_deaths ~ cause,
+    data = cleaned_data,
+    family = poisson(link = "log"),
+    seed = 888
+  )
+
+cause_of_death_canada_neg_binomial <-
+  stan_glm(
+    total_deaths ~ cause,
+    data = cleaned_data,
+    family = neg_binomial_2(link = "log"),
+    seed = 888
+  )
+
+modelsummary(
+  list(
+    "Poisson" = cause_of_death_canada_poisson,
+    "Negative binomial" = cause_of_death_canada_neg_binomial
+  ),
+)
+```
 
 ## Posterior predictive check
+@fig-2, @fig-3 indicate that negative binomial approach is a better choice for this circumstance. But it’s too early to draw conclusions from the figures alone.
+```{r echo = FALSE}
+#| label: fig-2
+#| warning: false
+#| fig-cap: "(a) Poisson model"
+pp_check(cause_of_death_canada_poisson) +
+  theme(legend.position = "bottom") +
+  theme_classic()
+```
+
+```{r echo = FALSE}
+#| label: fig-3
+#| warning: false
+#| fig-cap: "(b) Negative binomial model"
+pp_check(cause_of_death_canada_neg_binomial) +
+  theme(legend.position = "bottom") +
+  theme_classic()
+```
 There are many observations with a pareto_k > 0.7.
@@
 ## Resampling
-Following this logic, LOOCV(Leave-one-out cross-validation) will not be trustable anymore if most of the Pareto's K is higher than 0.7, although it can usually get a more accurate prediction.
+Following this logic, LOOCV(Leave-one-out cross-validation) will not be trustworthy anymore if most of the Pareto's K is higher than 0.7, although it can usually get a more accurate prediction.
 With this many problematic observations, K-fold cross-validation with argument 'K=10' to perform 10-fold cross-validation, that the data used for each training becomes closer to the entire data set, and the deviation decreases.
 
-It is less computationally intensive, and more efficient when processing large data sets.
-So K-fold will be a more suitable resampling method in this study.
+It is less computationally intensive, and more efficient when processing large data sets. So K-fold will be a more suitable resampling method in this study.
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+#| label: tbl-4
+#| tbl-cap: "K-fold cross-validation summary"
+poisson <- kfold(cause_of_death_canada_poisson, k = 10, cores = 1)
+neg_binomial <- kfold(cause_of_death_canada_neg_binomial, k = 10, cores = 1)
 
-The information provided in the table allows us to compare the relative performance of the negative binomial distribution model and the Poisson distribution model.
+comparison_results <- loo_compare(poisson, neg_binomial)
+rownames(comparison_results) <- c("neg_binomial", "poisson")
+
+kable(as.data.frame(comparison_results), booktabs = TRUE) %>%
+  kable_styling(font_size = 8, full_width = F)%>%
+  row_spec(0, bold = T)
+```
+The information provided in the @tbl-4 allows us to compare the relative performance of the negative binomial model and the Poisson model.
 The result mainly relies on the “elpd_diff” and “elpd_kfold” values of the two models, namely "expected log pointwise predictive density difference" and "ELPD for K-fold cross-validation".
@@
 
-From table, we can see that the negative binomial distribution model has higher “elpd_diff” and “elpd_kfold” values and also has a smaller standard deviation.
-This shows that the negative binomial distribution model has better predictive performance than the Poisson distribution model in cross-validation.
-Therefore, we can infer that the negative binomial distribution model is a better fit for the overall number of deaths by cause, in Canada, from 2000 to 2022.
+From table, we can see that the negative binomial model has higher “elpd_diff” and “elpd_kfold” values and also has a smaller standard deviation.
+This shows that the negative binomial model has better predictive performance than the Poisson model in cross-validation.
+Therefore, we can infer that the negative binomial model is a better fit for the overall number of deaths by cause, in Canada, from 2000 to 2022.
 
 \newpage
-# Discussion：
+# Discussion
 Plots of time-series changes in the number of deaths in Canada due to different factors are provided in the visualisation results of this study.
 The excellence of two forecasting models for mortality is analysed in comparison.
+
 ## Finding
-The main focus of this document revolves around a comparison of the performance of two statistical models - the negative binomial distribution model and the Poisson distribution model - in predicting mortality in Canada.
-Through detailed data analysis, we find that the negative binomial distribution model outperforms the Poisson distribution model in terms of both the expected log-point-by-point predicted density difference (elpd_diff) and the k-fold cross-validated ELPD value (elpd_kfold).
-In addition, the negative binomial distribution model has a smaller standard deviation, suggesting that its predictions are more stable.
-This finding provides a new perspective that negative binomial distribution models may be more appropriate for describing and predicting mortality in Canada.
+The main focus of this document revolves around a comparison of the performance of two statistical models - the negative binomial model and the Poisson model - in analyzing mortality in Canada.
+Through detailed data analysis, we find that the negative binomial model outperforms the Poisson model in terms of both the expected log-point-by-point predicted density difference(elpd_diff) and the K-fold cross-validated ELPD value(elpd_kfold).
+In addition, the negative binomial model has a smaller standard deviation, suggesting that its predictions are more stable.
+This finding provides a new perspective that negative binomial regression may be more appropriate for describing and predicting mortality in Canada.
 
 ## Economic Impact Insights
-From the perspective of economic impact, accurate mortality prediction is an important reference value for a number of industries, including insurance, pension, and healthcare.
-The superiority of the negative binomial distribution model means that we can predict future mortality rates more accurately, thus helping these industries to make more rational economic decisions, such as setting insurance premium rates and planning medical resources.
-In addition, for governments, accurate mortality prediction can also help to formulate more effective social security policies and reduce financial pressure.
-In terms of insight, the reason why the negative binomial distribution model can achieve better prediction results may be related to its ability to better handle the discrete and over-discrete nature of the data.
+From the perspective of economic impact, accurate mortality prediction is an important reference value for a number of industries, including insurance, pension, and healthcare[@1].
+The superiority of the negative binomial regression means that we can predict future mortality rates more accurately, thus helping these industries to make more rational economic decisions, such as setting insurance premium rates and planning medical resources.
+In addition, for governments, accurate mortality prediction can also help to formulate more effective social security policies and reduce financial pressure[@3].
+In terms of insight, the reason why the negative binomial model can achieve better prediction results may be related to its ability to better handle the discrete and over-discrete nature of the data.
 This suggests that we need to pay more attention to the characteristics of the data and choose more appropriate statistical models when dealing with similar problems.
@@
 In terms of social impact, accurate mortality forecasts help the public to better understand the trends in demographic and health conditions, and thus make more informed life decisions.
-For example, the public can adjust their health management and retirement planning based on the prediction results.
+For example, the public can adjust their health management and retirement planning based on the prediction results[@2].
 In terms of technical implications, the findings in this paper provide new ideas for the application of statistical models in the field of mortality prediction.
-In the future, we can further explore and optimise the negative binomial distribution model, or find other more suitable models to improve the prediction accuracy and stability.
+In the future, we can further explore and optimize the negative binomial distribution model, or find other more suitable models to improve the prediction accuracy and stability.
 
 ## Weakness and Future Research Directions
-Despite the superiority of the negative binomial distribution model in predicting mortality in Alberta, there are still some shortcomings.
+Despite the superiority of the negative binomial regression in predicting mortality in Canada, there are still some shortcomings.
 For example, only two models were compared in this paper and other possible models were not considered; furthermore, the parameters of the models were not analysed in detail to identify the key factors affecting the prediction results.
@@
 \newpage
-# Conclusion
+# Reference

### 2024-03-17 outputs/paper.qmd
sort out

@@
 ## Raw data
-First, the original dataset file of "Leading causes of death", named "1310039401-eng", was downloaded from Statistics Canada.
+First, the original dataset file of "Leading causes of death", called "1310039401-eng", was downloaded from Statistics Canada and renamed as "raw_data".
 Five key variables are included.
@@
 #| tbl-cap: "Modeling the most prevalent cause of deaths in Canada, 2000-2022"
-cleaned_data <- read.csv("cleaned_data")
+cleaned_data <- read.csv("inputs/cleaned_data.csv")
 cleaned_data <- cleaned_data[, -1]

### 2024-03-17 scripts/02-data_cleaning.R

@@ -2,7 +2,7 @@
 library(dplyr)
 
 # Read the CSV file as lines
-lines <- readLines("inputs/1310039401-eng.csv")
+lines <- readLines("inputs/raw_data.csv")
 
 # Skip specified lines
 lines_to_keep <- lines[-c(1:11, 13:15, 118:152)]

### 2024-03-17 outputs/paper.Rmd
1

--- 2024-03-17 outputs/paper.qmd
@@
 ---
+
 ```{r setup, include=FALSE}
@@
 ```
+
 \newpage
+
 # Introduction
+
 Studying public health mortality rates is of great significance to policy makers, healthcare professionals, and society as a whole.
@@
 \newpage
+
 # Data：
+
 This section seeks to provide a comprehensive understanding of the dataset used in our analysis.
@@
 ## Source
+
 Our study uses a dataset from Statistics Canada[@statcan] and focuses on mortality trends in Canada from 2000 to 2022, categorized by age, sex, and cause of death.
@@
 ### Measurement
+
 Using Poisson and negative binomial distributions as link functions, the data were fitted and the results of the two distributions were compared to select the best link function.
@@
 ## Raw data
-First, the original dataset file of "Leading causes of death", named "1310039401-eng", was downloaded from Statistics Canada.
+
+First, the original dataset file of "Leading causes of death", called "1310039401-eng", was downloaded from Statistics Canada and renamed as "raw_data".
 Five key variables are included.
@@
 ## Data cleaning
+
 In order to understand the data more intuitively, data cleansing is necessary.
@@
 ### Preview
+
 ```{r echo = FALSE}
@@
 ### Trend
-```{r echo = FALSE}
+
+```{r echo = FALSE, fig.width = 10, fig.height = 5}
 #| label: fig-1
@@
 \newpage
+
 # Result
+
 ```{r echo = FALSE}
@@
 ## Model fitting
+
 The Poisson distribution describes the probability distribution of the number of events that occur within a fixed unit of time or space.
@@
 ## Estimate
+
 Implementing Poisson and negative binomial regression, simultaneously.
@@
 According to the @tbl-3, the estimates are similar, so it is necessary to use a posterior predictive check in addition.
+
 ```{r  echo=FALSE, warning = FALSE}
@@
 ## Posterior predictive check
-@fig-2, @fig-3 indicate that negative binomial approach is a better choice for this circumstance. But it’s too early to draw conclusions from the figures alone.
-```{r echo = FALSE}
+
+@fig-2 indicates that negative binomial approach is a better choice for this circumstance. But it’s too early to draw conclusions from the figures alone.
+
+```{r echo = FALSE, warning = FALSE}
 #| label: fig-2
-#| warning: false
-#| fig-cap: "(a) Poisson model"
+#| fig-cap: "Fitting data into different models"
+#| fig-subcap: ["Poisson model", "Negative binomial model"]
+#| layout-ncol: 2
 pp_check(cause_of_death_canada_poisson) +
+  theme(legend.position = "bottom") +
+  theme_classic()
+
+pp_check(cause_of_death_canada_neg_binomial) +
   theme(legend.position = "bottom") +
@@
 
-```{r echo = FALSE}
-#| label: fig-3
-#| warning: false
-#| fig-cap: "(b) Negative binomial model"
-pp_check(cause_of_death_canada_neg_binomial) +
-  theme(legend.position = "bottom") +
-  theme_classic()
-```
-There are many observations with a pareto_k > 0.7.
+There are many observations with a pareto_k \> 0.7.
 Large Pareto's K means the model posterior would be too different if one data point is being removed.
@@
 ## Resampling
+
 Following this logic, LOOCV(Leave-one-out cross-validation) will not be trustworthy anymore if most of the Pareto's K is higher than 0.7, although it can usually get a more accurate prediction.
@@
 It is less computationally intensive, and more efficient when processing large data sets. So K-fold will be a more suitable resampling method in this study.
+
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
@@
 ```
+
 The information provided in the @tbl-4 allows us to compare the relative performance of the negative binomial model and the Poisson model.
@@
 \newpage
+
 # Discussion
+
 Plots of time-series changes in the number of deaths in Canada due to different factors are provided in the visualisation results of this study.
@@
 ## Finding
+
 The main focus of this document revolves around a comparison of the performance of two statistical models - the negative binomial model and the Poisson model - in analyzing mortality in Canada.
@@
 ## Economic Impact Insights
+
 From the perspective of economic impact, accurate mortality prediction is an important reference value for a number of industries, including insurance, pension, and healthcare[@1].
@@
 ## Societal and Technological Influences
+
 In terms of social impact, accurate mortality forecasts help the public to better understand the trends in demographic and health conditions, and thus make more informed life decisions.
@@
 ## Weakness and Future Research Directions
+
 Despite the superiority of the negative binomial regression in predicting mortality in Canada, there are still some shortcomings.
@@
 \newpage
+
 # Reference

### 2024-03-17 outputs/paper.qmd
1

--- 2024-03-17 outputs/paper.qmd
@@
 ```
-
 Preview the processed clean data first, which is very helpful for understanding the overall data.
@@
 ### Trend
-```{r echo = FALSE}
+```{r echo = FALSE, fig.width = 10, fig.height = 5}
 #| label: fig-1
@@
 ```
-
 @fig-1 shows how many people died from different causes each year in Canada from 2000 to around 2022.
@@
 ## Posterior predictive check
-@fig-2, @fig-3 indicate that negative binomial approach is a better choice for this circumstance. But it’s too early to draw conclusions from the figures alone.
-```{r echo = FALSE}
+@fig-2 indicates that negative binomial approach is a better choice for this circumstance. But it’s too early to draw conclusions from the figures alone.
+```{r echo = FALSE, warning = FALSE}
 #| label: fig-2
-#| warning: false
-#| fig-cap: "(a) Poisson model"
+#| fig-cap: "Fitting data into different models"
+#| fig-subcap: ["Poisson model", "Negative binomial model"]
+#| layout-ncol: 2
 pp_check(cause_of_death_canada_poisson) +
@@
   theme_classic()
-```
 
-```{r echo = FALSE}
-#| label: fig-3
-#| warning: false
-#| fig-cap: "(b) Negative binomial model"
 pp_check(cause_of_death_canada_neg_binomial) +

### 2024-03-17 outputs/paper.Rmd
1

@@
 #| tbl-cap: "Modeling the most prevalent cause of deaths in Canada, 2000-2022"
-cleaned_data <- read.csv("cleaned_data")
+cleaned_data <- read.csv(here::here("cleaned_data"))
 cleaned_data <- cleaned_data[, -1]

### 2024-03-17 outputs/paper.qmd
1

@@
 #| tbl-cap: "Modeling the most prevalent cause of deaths in Canada, 2000-2022"
-cleaned_data <- read.csv("cleaned_data")
+cleaned_data <- read.csv(here::here("inputs/cleaned_data"))
 cleaned_data <- cleaned_data[, -1]
